### 2021-04-28 assertions/doc/thesis-bernhardt/chapters/0_2-abstract.tex
Conclusion
Revision

\chapter*{Abstract}
Modern blockchains provide a platform for building applications with Smart Contracts. Because transactions are processed at every node in the network, the quantity and complexity of computation, which can be performed by Smart Contracts, is limited. Furthermore, complex contracts can generate high transaction costs for the user. To overcome this limitation and increase cost efficiency, blockchain-based applications could move computations to an off-chain component and only process the result on the blockchain.
This thesis proposes a processing model, which engages the validators of the network in a distributed effort to verify a submitted off-chain computation. Based on a specification of properties, the validators are challenged to find a counterexample by asserting these properties for independent subsets of the result. This reduces the computational effort on the blockchain to the verification of the result, instead of the computation of the result itself. Due to the distributed scheme, the computations per node are decreased even further. As a further contribution, this thesis defines a domain-specific language to express explicit assertions for off-chain computations. It describes the model and a proof-of-concept implementation of a processing pipeline, which transforms and compiles these assertions into Smart Contracts for the target platform Tezos. These contracts conduct a point-wise search for a counterexample and are called during the verification process.

%Modern blockchains support building own applications with Smart Contracts.
%because transactions are processed by every node, the amount and complexity of computation which can be performed by such SC is limited. %Furthermore, complex SC can cause high transaction costs for the user.
%To overcome this limitation and increase efficiency, applications could move computations to an off-chain component and submit the results to the blockchain. 
%This thesis proposes a processing model which engages the validators of the network to a distributed validation of such a submitted result. It challenges them to find counterexamples by performing an assertion check for independent subsets of the submitted result.
% This reduces the computational efforts on the chain to the verification of the result instead of computing the result itself.
% due to the distributed approach, the computation per validator and node is reduced even further

% Furthermore, this thesis defines a domain-specific language to express explicit assertions over off-chain computations. It also provides a pipeline description for a toolchain, which transforms and compiles these assertions into SC searching for counterexamples. These contracts are then used in the distributed effort to validate the result.
% this toolchain was partially implemented and provided with this thesis.

% Tezos!

>>>>

\chapter*{Abstract}
Modern blockchains provide a platform for building sophisticated applications with Smart Contracts. Because transactions are processed at every node in the network, the quantity and complexity of computations, which can be performed by Smart Contracts, are limited. Furthermore, complex contracts can generate high transaction costs for the user. To overcome this limitation and improve cost efficiency, blockchain-based applications could move computations to an off-chain component and only process the result on the blockchain. This thesis proposes a processing model, which engages the validators of the network in a distributed effort to verify a submitted off-chain computation. Based on a specification of properties, the validators are challenged to find a counterexample by asserting these properties for independent subsets of the result. This reduces the computational effort on the blockchain to the verification of the result, rather than the computation of the result itself. Moreover, by implementing a distributed scheme, each node  only has to process a fraction of the verification. As a further contribution, this thesis defines a domain-specific language to express explicit assertions for off-chain computations and specifies a processing pipeline to transform and compile these assertions into Smart Contracts. These contracts conduct a point-wise search for a counterexample and are called during the verification process. A proof-of-concept implementation of this pipeline is provided and described for the target platform Tezos.

### 2021-04-28 assertions/doc/thesis-bernhardt/chapters/2-use_case_analysis.tex
Revision

\section{Restriction of use-cases}\label{sec:restrict}\todo{Better title? -> conclusion?}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking. Additionally, they require a distinct translation to target code. This increases the complexity and requirements for an implementation of the toolchain and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

@@

\figref{fig:use_case_cost} shows the reported gas consumption of transactions calling corresponding dummy contracts in Michelson (cf. Appendix \ref{apx:cost_analysis_contract}) on the Delphinet test network \footnote{\url{https://tezos.gitlab.io/introduction/test_networks.html\#delphinet}}. As to be expected, the gradient of the cost function in the case of \texttt{Bar.bar} is higher due to additional serialization and parsing costs, namely $\frac{dx}{dy} = 16$ compared to $\frac{dx}{dy} = 6$ for \texttt{Foo.foo}.
\begin{figure}[h]
\centering
\includegraphics[width=0.9\linewidth]{figures/2-use_cases/cost_analysis}
\caption{Gas consumption of transactions calling \texttt{Foo} and \texttt{Bar} for increasing input sizes}
\label{fig:use_case_cost}
%\vspace{128in}
\end{figure}

>>>>

\section{Restriction of use-cases}\label{sec:restrict}\todo{Better title?}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking. Additionally, they require a distinct translation to target code. This increases the complexity and requirements for an implementation of the toolchain and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

@@

\figref{fig:use_case_cost} shows the reported gas consumption of transactions calling corresponding dummy contracts in Michelson (cf. Appendix \ref{apx:cost_analysis_contract}) on the Delphinet test network \footnote{\url{https://tezos.gitlab.io/introduction/test_networks.html\#delphinet}}. As to be expected, the gradient of the cost function in the case of \texttt{Bar.bar} is higher due to additional serialization and parsing costs, namely $\frac{dx}{dy} = 16$ compared to $\frac{dx}{dy} = 6$ for \texttt{Foo.foo}.
\begin{figure}[h]
\centering
\includegraphics[width=0.9\linewidth]{figures/2-use_cases/cost_analysis}
\caption{Gas consumption of contracts \texttt{Foo} and \texttt{Bar} for increasing inputs}
\label{fig:use_case_cost}
%\vspace{128in}
\end{figure}

### 2021-04-28 assertions/doc/thesis-bernhardt/chapters/5-offline_design_tezos.tex
Revision

\subsection{Representing counterexamples}\label{sec:counterexample}
It remains to be specified, how counterexamples are represented within the network. The interaction scheme in \figref{fig:interaction_modular} suggests, that validators publish a counterexample to the network in case of a failed assertion. For this, the validators have to obtain the randomly generated values from within the execution context of the VM. With the \texttt{failwith} instruction, the top element of the stack can be exposed, allowing to return an interpretable expression from the execution context. The generated values can thus be returned as a single value, a (nested) pair or a list. To allow other validators to check the assertion with a specific set of values, a modified version of the assertion code must be available on the blockchain. Instead of generating random values, this version should take these values as input parameters. Hence, the validator creates a normal transaction invoking this code with the values returned by the \texttt{failwith} instruction and publishes it to the network. The counterexample is then validated like any other transaction on the blockchain.

To implement this scheme, the compiler has to generate one more contract per assertion. In addition, the assertion contract needs to know about the address of the counterexample contract and return the address together with the generated values. This requires a adjustment of the storage type of the assertion contract and changes the order of origination to:
\begin{enumerate}
\itemsep-0.5em
\item \texttt{Origination(parent)}, \texttt{Origination(counterexample)}
\item \texttt{Origination(assertion, address counterexample)}
\item \texttt{Origination(controller, address [assertion], address parent)} 
\end{enumerate}

\subsection{Multi-slot validators}
The proposed implementation for Tezos assumes that all 32 validators of the current block cycle execute a determined number of test runs. In fact, as mentioned in the section describing Tezos' proof-of-stake mechanism, a validator may have been assigned several endorsement slots per cycle. Since these validators only execute the assertion once, the test runs for the other endorsement slots are neglected. As a result, the error rate of the assertion check is unpredictable and may exceed the configured threshold. This issue should be addressed by the protocol design --- multi-slot validators should execute the assertion transaction for each of their endorsement slot.

>>>>

\subsection{Representing counterexamples}\label{sec:counterexample}
It remains to be specified, how counterexamples are represented within the network. The interaction scheme in \figref{fig:interaction_modular} suggests, that validators publish a counterexample to the network in case of a failed assertion. For this, the validators have to obtain the randomly generated values from within the execution context of the VM. With the \texttt{failwith} instruction, the top element of the stack is exposed, allowing to return an interpretable expression from the execution context. The generated values can thus be returned as a single value, a (nested) pair or a list. To allow other validators to check the assertion with a specific set of values, a modified version of the assertion code must be available on the blockchain. Instead of generating random values, this version should take these values as input parameters. Hence, the validator could create a normal transaction invoking this code with the values returned by the \texttt{failwith} instruction and publish it to the network. The counterexample is then validated like any other transaction on the blockchain.

To implement this scheme, the compiler has to generate one more contract per assertion. In addition, the assertion contract needs to know about the address of the counterexample contract, and return the address together with the generated values in case of a failure. This requires an adjustment of the storage type of the assertion contract and changes the order of origination to:
\begin{enumerate}
\itemsep-0.5em
\item \texttt{Origination(parent)}, \texttt{Origination(counterexample)}
\item \texttt{Origination(assertion, address counterexample)}
\item \texttt{Origination(controller, address [assertion], address parent)} 
\end{enumerate}

In order to reject a transaction submitting off-chain computations in case of a published counterexample, however, the network needs to be able to relate a counterexample to the respective transaction.

\subsection{Multi-slot validators}
The proposed implementation assumes that all 32 validators of the current block cycle execute a determined number of test runs. In fact, as mentioned in the section describing Tezos' proof-of-stake mechanism, a validator may have been assigned several endorsement slots per cycle. Since these validators only execute the assertion once, the test runs for the other endorsement slots are neglected. As a result, the error rate of the assertion check is not deterministic and may exceed the configured threshold. This issue should be addressed by the protocol design --- multi-slot validators should execute the assertion check for each of their endorsement slots.

### 2021-04-28 assertions/doc/thesis-bernhardt/chapters/6-conclusion.tex
Revision

For the proposed architecture, the conducted costs analysis shows high transaction fees for the submission of off-chain computations,  due to high verification costs. Although the verification might still be more efficient than the execution itself, it can be assumed that the cost reductions are narrowed, or even equalised in cases of moderate complexity of off-chain computations. Inter-contract communication was identified as a main driving factor. To lower these costs and approach usability on production systems, the thesis provided possible improvements and modifications of the architecture.

If it is possible to make the verification more efficient, this approach provides a simple solution for increasing of scalability of blockchain-based applications. Assertions can be specified in an expressive and familiar way and the processing model only requires moderate adaptions to the blockchain protocol.

@@

Besides the completion of the proof-of-concept implementation, the assertion language can be extended with features, such as variables and user-defined functions, to improve (re-)usability. Furthermore, an interesting extension is the support for properties formulated with existential quantification, as this allows to verify off-chain computations through distributed validation instead of refutation.

>>>>

For the proposed architecture, the conducted costs analysis shows high transaction fees for the submission of off-chain computations,  due to high verification costs. Although the verification might still be more efficient than the execution itself, it can be assumed that the cost reductions are equalised in cases of moderate complexity of off-chain computations. Inter-contract communication was identified as a main driving factor of the verification cost. To lower these costs and approach usability on production systems, the thesis provided possible improvements and modifications of the architecture.

In case the verification cost can be lowered to a feasible threshold, this approach provides a simple solution for increasing the scalability of blockchain-based applications. Assertions can be specified in an expressive and familiar way, and the processing model only requires moderate adaptions to the blockchain protocol.

@@

Besides the completion of the proof-of-concept implementation, the assertion language can be extended with features, such as variables and user-defined functions, to improve (re-)usability. Furthermore, an interesting extension is the support for properties formulated with existential quantification, as this allows to verify off-chain computations by distributed validation in addition to distributed refutation.
